Robustly decorrelating errors with mixed quantum gates

### 2020-09-11 decorrelation/resubmission/response_to_referees.txt

First Referee,
Thank you for the helpful suggestions. In line with your comment, we have added the following text to Secs. IIIB3 and IVB1 with Fig. 6 for readers not familiar with sparse modeling.
""

Second Referee,
Thank you for your report.

While randomized compiling is well known in the field and affects the same outcome as our strategy, it is important to note that our technique works at the hardware level, rather than the compilation level. Thus our routine does not require a precompilation step for each circuit. Additionally, as you point out, our routine has been cast as a convex optimization routine. Because of the way we have formulated our problem, we are able to augment the mixed quantum gate to have properties besides just twirling away the error. We have summarized this in the text now as:
""

We will now respond to your five bullet points.

The experiment we chose was meant to demonstrate the reduced coherent error of the channel. While it's true that any mixing of the unitaries will outperform any of the constituent unitaries as measured by RB number, the demonstration was meant to show a reduction in the coherence of the error. This is the result mentioned by Ball et al., and we have revised the text to emphasize this. Had we chosen different weights than those produced by the convex optimizer, the error bars would have been larger.

This is a great suggestion, although the presentation might be more cluttered. However, the first author no longer has access to the hardware at Rigetti Computing.

While it is true that RB randomizes errors and our technique also randomizes errors, randomized benchmarking still show cases the property of the resulting channel that we want. Namely, that the errors have minimal coherent error. Using an algorithm like QAOA would be challenging to demonstrate a decrease in the coherent error of the resulting channel.

The two-qubit experiment is substantially harder to do. Two qubit infidelities are sufficiently high that even repeating the randomized benchmarking experiment would have very poor signal to noise, since an average clifford gate has a large number of two qubit gates. While we agree that a two-qubit demonstration would be a good addition to the manuscript, it is also the case the author no longer has access to the hardware.

The control errors need to be known to you - you can accomplish this through something like tomography. You might object - if I know my control errors, I will just fix them! It could be the case that you don't have enough samples in your DAC to do this, or that the coherent errors come from unknown interactions that are out of your control. Moreover, in the context of drift as in superconducting platforms, it is the case that you can pick out a family of controls that will all drift, but drift in different ways such that their average channel performs in a better way than any individual channel. In our experiment we have introduced the miscalibrations intentionally, since the qubits were coherence limited. We have clarified this in the text as a summary of the protocol:

------

Kevin's Edits:

### 2020-09-11 decorrelation/resubmission/response_to_referees.txt

Here I have a comment for readability. The presentation in Secs. IIIB3 and IVB1 with Fig. 6 may be difficult to understand, especially for readers not familiar with sparse modeling. I recommend the improvement of the parts.

We thank the referee for their positive assessment of our work, as well as for alerting us to potentially unclear sections of the manuscript. In line with these comments, we have added the following text to Secs. IIIB3 and IVB1 and modified Fig. 6 to improve the clarity for readers not familiar with sparse optimization.

Here, $\lambda \geq 0$ is a tunable parameter that can be used to control the degree of sparsity in the solution. If $\lambda$ is set to zero, the inner program reduces to Eq.~\eqref{eq:minimization} and the outer minimization is redundant. If $\lambda$ is very large, each inner minimization problem is minimized for an MQG with a single control, and the outer minimization simply selects the control which has the error generator of smallest norm. While this gives a maximally sparse solution, it doesn't allow for any cancellation of coherent errors.

In general, increasing $\lambda$ decreases the number of controls with large weights, and by removing controls with weights below a threshold probability we can increase the sparsity of the solution. Truncating in this way will introduce error in the generator-exactness of the solution. For larger $\lambda$ the required threshold probability decreases, and so the error introduced from truncation can be made smaller. However, larger $\lambda$ will also decrease the likelihood of the optimizer returning a generator-exact solution, and so in this way, tuning $\lambda$ allows us to trade between different sources of error when constructing sparse MQGs.

@@

Randomized compiling is technique known to the QC community for quite some time now and therefore there is nothing particularly novel about the technique itself. What appears to be the highlight of the paper is the convex algorithm that the authors used to choose random gates and the experimental demonstration. While these are good steps toward a nice direction, I feel the demonstrations are not sufficiently convincing, particularly in regard to the experimental part (which the authors themselves acknowledge as being "contrived"). Here are my main concerns:

While the referee is correct that our approach is superficially similar to randomized compiling, we wish to emphasize that our technique is intended as an alternative approach to solving the problem of coherent errors in quantum information processors. Our method is particularly well suited for experimental implementation, as the mixed unitary gates can be implemented at the level of individual logic operations using the FPGA control hardware. Contrast this against the randomized compiling approach, which requires wholesale recompilation of the circuit every time it is run. Importantly, though, we make no claim that this broad approach is novel – in fact, we frequently cite prior foundational work in this area. Instead, our manuscript extends and expands the scope of this prior work, formulating the problem as an explicit convex optimization, incorporating robustness to uncertainty, and, critically, providing the first experimental demonstration of this technique on quantum information processing hardware. We have added the following to the text to clarify this point:

MQGs are an alternative approach to solving this problem. Because they are implemented by randomly drawing from different gate realizations, they only require randomness at run-time, allowing them to be easily implemented in hardware. Moreover, when the construction of MQGs is posed as a numerical optimization problem, numerous useful modifications can be made to augment their performance.

### 2020-10-07 decorrelation/mixed_unitaries.tex
anthony made some changes before submitting

Randomized protocols have a long history of outperforming their deterministic counterparts\cite{Viola2005, Santos2006}. In circuit models, Pauli Frame Randomization \cite{Wallman2016, 1803.01818, dahlen2018experimental} has been shown to reduce coherent errors by introducing randomness in a compilation step. MQGs are an alternative approach to solving this problem. Because they are implemented by randomly drawing from different gate realizations, they only require randomness at run-time, allowing them to be easily implemented in hardware. Moreover, when the construction of MQGs is posed as a numerical optimization problem, numerous useful modifications can be made to augment their performance.

In this article, we extend the work of Campbell and Hastings to numerically optimized quantum gates and show that the advantages of their approach can be made robust to drift and model uncertainty. We provide an explicit series of convex programs able to efficiently compute MQGs with a variety of useful properties. We demonstrate our results on a superconducting testbed device at Rigetti Computing.  In a simple experiment based on randomized benchmarking circuits, MQGs demonstrate a marked improvement in error rates and a reduced variance in circuit outcome probabilities, consistent with a significant reduction in the coherence of the error \cite{Ball2016}. We further apply our methods in simulation, constructing both single- and two-qubit mixed unitary controls that are robust to drift and uncertainty in the control parameters.

@@

We can additionally generalize the geometric criterion for guaranteeing the existence of generator-exact MQGs. To construct a robust, generator-exact MQG, there must exist a set of weights, $\{\weight_i\}$, so that $\sum_i \weight_i \genmat_i = \sum_i \weight_i \genmat_{i,k} = 0$. This is equivalent to demanding the origin lie in the convex hull of the generators augmented with their derivatives:
\begin{equation}
\0 \in \mathsf{conv}\left( \left\{
\genmat_i \bigoplus_k \genmat_{i,k}
\right\}_i	 \right).
\end{equation}
We illustrate this in Fig.~\ref{fig:vectorspace2}.

We call a gate constructed in this way a \emph{robust} generator-exact MQG. These techniques can be easily adapted to construct robust Pauli-exact MQGs, where vectors of the off-diagonal elements of the error maps take the role of the error generators used above.

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{vectorspace3.png}
  \caption{\textbf{Geometric condition for robustness:}   A target unitary gate can be implemented a number of ways, each with a different effective error generator, $\genmat_i$. a) The error generators are shown as elements of a vector space (solid, black arrows). Also shown are their derivatives $\genmat_{i,k}$ (dashed, grey arrows) with respect to a model parameter, $\delta_k$. As this parameter drifts, the generators may no longer cancel. b) To construct a robust, generator-exact MQG, there must exist a set of weights, $\{\weight_i\}$, so that $\sum_i \weight_i \genmat_i = \sum_i \weight_i \genmat_{i,k} = 0$. This is equivalent to demanding that the origin lie in the convex hull $\mathsf{conv}(\{\genmat_i \bigoplus_k \genmat_{i,k} \}_i)$. The gradient arrows indicate that this space is higher-dimensional. }
  \label{fig:vectorspace2}
\end{figure}

@@

\subsection{Experimental implementation } % (fold)
\label{sub:experimental}
\noindent We implemented our methods using the 19Q-Acorn superconducting transmon processor at Rigetti Computing. Comprehensive characterization of this device can be found in \cite{1712.05771}. While this device consisted of 20 qubits with fixed couplings, we used only qubit \#8 for our demonstration.

@@

We note that, while this example is somewhat contrived (the calibrated gate is clearly the best), it nonetheless succeeds in demonstrating that the MQGs are capable of outperforming the individual gates of which they are composed. In future work, we hope to apply this technique to two-qubit gates, which are significantly more difficult to tune up in general and are more likely to possess lingering coherent errors. Tomography or modeling can then be used to construct the process matrix estimates required to build a mixed gate.

@@

In this example, we find more modest improvements to performance, as shown in Fig. \ref{fig:2MQG}. There are now four free parameters to optimize over, and the uncontrolled entangling interaction means that there is little room for variation in the controls. Nonetheless, using an MQG improves performance over any of the constituent controls over a wide-range of detunings, and for all values of the detunings we see that the robust generator-exact MQG performs as well or better than generator-exact MQG.

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{2q_robustness.pdf}
  \caption{\textbf{Two-qubit robust generator-exact MQG} Numerical results showing the diamond distance errors for a generator-exact MQG (dashed blue), a robust generator-exact MQG (solid orange), and a large set of bare unitary controls (thin solid grey) as a function of a spurious detuning (quantified by the ratio of the detuning to the maximum control amplitude). It is intended to model a pair of qubits interacting via resonant exchange. Both MQGs can be seen to outperform the bare unitary control over a wide range of detunings. The robust generator-exact  MQG always outperforms the generator-exact MQG, and both MQGs outperform any of the bare unitary controls over a wide range of detunings. Similar results are obtained for drift in the qubit control amplitude.}
  \label{fig:2MQG}
\end{figure}

>>>>

In this article, we extend the work of Campbell and Hastings to numerically optimized quantum gates and show that the advantages of their approach can be made robust to drift and model uncertainty. We provide an explicit series of convex programs able to efficiently compute MQGs with a variety of useful properties. We demonstrate our results on a superconducting testbed device at Rigetti Computing. In a simple experiment based on randomized benchmarking circuits, MQGs demonstrate a marked improvement in error rates and a reduced variance in circuit outcome probabilities, consistent with a significant reduction in the coherence of the error \cite{Ball2016}. We further apply our methods in simulation, constructing both single- and two-qubit mixed unitary controls that are robust to drift and uncertainty in the control parameters. These robust MQGs are insensitive to calibration errors, but do not require the temporal overhead of coherent control techniques\cite{1203.6392}.

Randomized protocols have a long history of outperforming their deterministic counterparts\cite{Viola2005, Santos2006}. In circuit models, Pauli Frame Randomization \cite{Wallman2016, 1803.01818, dahlen2018experimental} has been shown to reduce coherent errors by introducing randomness in a compilation step. MQGs are an alternative approach to solving this problem. Because they are implemented by randomly drawing from different gate realizations, they only require randomness at run-time, allowing them to be easily implemented in hardware. Moreover, as we will discuss, when the construction of MQGs is posed as a numerical optimization problem, numerous useful modifications can be made to augment their performance.

@@

We can additionally generalize the geometric criterion for guaranteeing the existence of generator-exact MQGs. To construct a robust, generator-exact MQG, there must exist a set of weights, $\{\weight_i\}$, so that $\sum_i \weight_i \genmat_i = \sum_i \weight_i \genmat_{i,k} = 0$. This is equivalent to demanding the origin lie in the convex hull of the generators augmented with their derivatives:
\begin{equation}
\0 \in \mathsf{conv}\left( \left\{
\genmat_i \bigoplus_k \genmat_{i,k}
\right\}_i	 \right).
\end{equation}
We illustrate this in Fig.~\ref{fig:vectorspace2}. We call a gate constructed in this way a \emph{robust} generator-exact MQG. These techniques can be easily adapted to construct robust Pauli-exact MQGs, where vectors of the off-diagonal elements of the error maps take the role of the error generators used above. An important property of these robust MQGs is their relatively short implementation times, as compared with conventional coherent control techniques. To see this, consider two well-known robust optimal control sequences, BB1 and SK1.\cite{Wimperis1994, Brown2004}. Considering specifically pulse amplitude errors, BB1 utilizes a so-called broadband sequence which requires four rotations to mitigate calibration errirs. In the more general case of SK1 (and SKn) pulses, we consider an arbitrary error of magnitude $\epsilon$, that we wish to supress to order $n$. Already, for an SK1 pulse, this technique requires three rotation. Contrast this with the duration of an MQG, which is at worst the duration of the longest gate in the family.
\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{vectorspace3.png}
  \caption{\textbf{Geometric condition for robustness:}   A target unitary gate can be implemented a number of ways, each with a different effective error generator, $\genmat_i$. a) The error generators are shown as elements of a vector space (solid, black arrows). Also shown are their derivatives $\genmat_{i,k}$ (dashed, grey arrows) with respect to a model parameter, $\delta_k$. As this parameter drifts, the generators may no longer cancel. b) To construct a robust, generator-exact MQG, there must exist a set of weights, $\{\weight_i\}$, so that $\sum_i \weight_i \genmat_i = \sum_i \weight_i \genmat_{i,k} = 0$. This is equivalent to demanding that the origin lie in the convex hull $\mathsf{conv}(\{\genmat_i \bigoplus_k \genmat_{i,k} \}_i)$. The gradient arrows indicate that this space is higher-dimensional. }
  \label{fig:vectorspace2}
\end{figure}

@@

\subsection{Experimental implementation} % (fold)
\label{sub:experimental}
\noindent We implemented our methods using the 19Q-Acorn superconducting transmon processor at Rigetti Computing. Comprehensive characterization of this device can be found in \cite{1712.05771}. While this device consisted of 20 qubits with fixed couplings, we used only qubit \#8 for our demonstration.

@@

We note that, while this example is somewhat contrived (the calibrated gate is clearly the best), it nonetheless succeeds in demonstrating that the MQGs are capable of outperforming the individual gates of which they are composed and are able to be easily implemented on modern devices. In future work, we hope to apply this technique to two-qubit gates, which are significantly more difficult to tune up in general and are more likely to possess lingering coherent errors.  In this case, tomography or modeling can then be used to construct the process matrix estimates required to build a mixed gate. Even in the case that the errors cannot be learned tomographically or modeled accurately, the robust methods introduced in the previous section can be used to produce MQGs with improved performance.

@@

In this example, we find more modest improvements to performance, as shown in Fig. \ref{fig:2MQG}. There are now four free parameters to optimize over, and the uncontrolled entangling interaction means that there is little room for variation in the controls. Nonetheless, using an MQG improves performance over any of the constituent controls over a wide-range of detunings, and for all values of the detunings we see that the robust generator-exact MQG performs as well or better than generator-exact MQG. And, as discussed previously, the robustness comes without the temporal overhead of introducing more gates sequentially.
\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{2q_robustness.pdf}
  \caption{\textbf{Two-qubit robust generator-exact MQG} Numerical results showing the diamond distance errors for a generator-exact MQG (dashed blue), a robust generator-exact MQG (solid orange), and a large set of bare unitary controls (thin solid grey) as a function of a spurious detuning (quantified by the ratio of the detuning to the maximum control amplitude). It is intended to model a pair of qubits interacting via resonant exchange. Both MQGs can be seen to outperform the bare unitary control over a wide range of detunings. The robust generator-exact  MQG always outperforms the generator-exact MQG, and both MQGs outperform any of the bare unitary controls over a wide range of detunings. Similar results are obtained for drift in the qubit control amplitude. The insensitivity to $\epsilon$ here is reminiscent of the insensitivity to pulse amplitude in \cite{Wimperis1994}, however we emphasize that the robustness here is achieved without increasing the duration of the gate.}
  \label{fig:2MQG}
\end{figure}
